---
title: "MODELOS PREDICTIVOS (SECCION 9)"
output: html_document
date: "2024-04-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# CHAPTER 9: FORECASTING with Linear Regression

## I: Features engineering of the series components

### 1

```{r}
library(TSstudio)

data("USgas")

ts_plot(USgas,
        title = "US Monthly Natural Gas consumption",
        Ytitle = "Billion Cubic Feet",
        Xtitle = "Year")

ts_info(USgas)

```


```{r}
ts_decompose(USgas)
```

```{r}
USgas_df <- ts_to_prophet(USgas)
head(USgas_df)
USgas_df$trend <- 1:nrow(USgas_df)
```

### 2

```{r}
library(lubridate)

USgas_df$seasonal <- month(USgas_df$ds, label = T)
```

```{r}
head(USgas_df)
```

```{r}
h <- 12 # setting a testing partition length

train <- USgas_df[1:(nrow(USgas_df) - h), ]

test <- USgas_df[(nrow(USgas_df) - h + 1):nrow(USgas_df), ]
```

##  II: Modeling the series trend and seasonal components

### 1

```{r}
md_trend <- lm(y ~ trend, data = train)

summary(md_trend)
```

### 2

```{r}
train$yhat <- predict(md_trend, newdata = train)

test$yhat <- predict(md_trend, newdata = test)

library(plotly)
plot_lm <- function(data, train, test, title = NULL){
  p <- plot_ly(data = data, 
               x = ~ ds, 
               y = ~ y, 
               type = "scatter",
               mode = "line",
               name = "Actual") %>%
    add_lines(x =  ~ train$ds,
              y = ~ train$yhat,
              line = list(color = "red"),
              name = "Fitted") %>%
    add_lines(x =  ~ test$ds,
              y = ~ test$yhat,
              line = list(color = "green", dash = "dot", width = 3),
              name = "Forecasted") %>%
    layout(title = title,
           xaxis = list(title = ""),
           yaxis = list(title = "Billion Cubic Feet"),
           legend = list(x = 0.05, y = 0.95))
  return(p)
}
```

```{r}
plot_lm(data = USgas_df, 
        train = train, 
        test = test,
        title = "Predicting the Trend Component of the Series")
```

```{r}
mape_trend <- c(mean(abs(train$y - train$yhat) / train$y),
                mean(abs(test$y - test$yhat) / test$y))

mape_trend
```
### 3


```{r}
md_seasonal <- lm(y ~ seasonal, data = train)

summary(md_seasonal)
```

```{r}
train$yhat <- predict(md_seasonal, newdata = train)
test$yhat <- predict(md_seasonal, newdata = test)

plot_lm(data = USgas_df, 
        train = train, 
        test = test,
        title = "Predicting the Seasonal Component of the Series")
```

```{r}
mape_seasonal <- c(mean(abs(train$y - train$yhat) / train$y),
                   mean(abs(test$y - test$yhat) / test$y))

mape_seasonal
```


### 4


```{r}
md1 <- lm(y ~ seasonal + trend, data = train)

summary(md1)
```

```{r}
train$yhat <- predict(md1, newdata = train)
test$yhat <- predict(md1, newdata = test)


plot_lm(data = USgas_df, 
        train = train, 
        test = test,
        title = "Predicting the Seasonal Component of the Series")
```

```{r}
mape_md1 <- c(mean(abs(train$y - train$yhat) / train$y),
              mean(abs(test$y - test$yhat) / test$y))
mape_md1
```

### 5

```{r}
md2 <- lm(y ~ seasonal + trend + I(trend^2), data = train)

summary(md2)
```

```{r}
train$yhat <- predict(md2, newdata = train)
test$yhat <- predict(md2, newdata = test)


plot_lm(data = USgas_df, 
        train = train, 
        test = test,
        title = "Predicting the Seasonal Component of the Series")

mape_md2 <- c(mean(abs(train$y - train$yhat) / train$y),
              mean(abs(test$y - test$yhat) / test$y))

mape_md2
```

## III: The tslm function

### 1: Modeling single events and non-seasonal events

```{r}
USgas_split <- ts_split(USgas, sample.out = h)

train.ts <- USgas_split$train

test.ts <- USgas_split$test
```

```{r}
library(forecast)

md3 <- tslm(train.ts ~ season + trend + I(trend^2))

summary(md3)
```

```{r}
r <- which(USgas_df$ds == as.Date("2014-01-01"))
USgas_df$s_break <- ifelse(year(USgas_df$ds) >= 2010, 1, 0)
USgas_df$s_break[r] <- 1
md3 <- tslm(USgas ~ season + trend + I(trend^2) + s_break, data = USgas_df)
summary(md3)
```
### 2: The UKgrid series

```{r}
library(UKgrid)

UKdaily <- extract_grid(type = "data.frame",
                        columns = "ND",
                        aggregate = "daily")

head(UKdaily)
```
```{r}
ts_plot(UKdaily,
        title = "The UK National Demand for Electricity",
        Ytitle = "MW",
        Xtitle = "Year")
```


```{r}
ts_heatmap(UKdaily[which(year(UKdaily$TIMESTAMP) >= 2016),],
           title = "UK the Daily National Grid Demand Heatmap")
```
### 3: Preprocessing and feature engineering of the UKdaily series

```{r}
library(dplyr)
UKdaily <- UKdaily %>%
  mutate(wday = wday(TIMESTAMP, label = TRUE),
         month = month(TIMESTAMP, label = TRUE),
         lag365 = dplyr::lag(ND, 365)) %>%
  filter(!is.na(lag365)) %>%
  arrange(TIMESTAMP)
str(UKdaily)

```
### 4

```{r}
start_date <- min(UKdaily$TIMESTAMP)


UK_ts <- ts(UKdaily$ND, 
            start = c(year(start_date), yday(start_date)),
            frequency = 365)
```

```{r}
#ts_acf(UK_ts, lag.max = 365 * 4)
```
```{r}
h <-  365
UKpartitions <- ts_split(UK_ts, sample.out = h)
train_ts <- UKpartitions$train
test_ts <- UKpartitions$test

train_df <- UKdaily[1:(nrow(UKdaily) - h), ]
test_df <- UKdaily[(nrow(UKdaily) - h + 1):nrow(UKdaily), ]
```

## IV: Training and testing the forecasting model

### 1
```{r}
md_tslm1 <- tslm(train_ts ~ season + trend)
fc_tslm1 <- forecast(md_tslm1, h = h)

test_forecast(actual = UK_ts,
              forecast.obj = fc_tslm1,
              test = test_ts)
```
```{r}
accuracy(fc_tslm1, test_ts)
```
 
### 2

```{r}
md_tslm2 <- tslm(train_ts ~ season + trend + wday, data = train_df)
fc_tslm2 <- forecast(md_tslm2, h = h, newdata = test_df)

test_forecast(actual = UK_ts,
              forecast.obj = fc_tslm2,
              test = test_ts)
```


```{r}
accuracy(fc_tslm2, test_ts)
```

### 3

```{r}
md_tslm3 <- tslm(train_ts ~ season + trend + wday + month + lag365, data = train_df)
fc_tslm3 <- forecast(md_tslm3, h = h, newdata = test_df)

test_forecast(actual = UK_ts,
              forecast.obj = fc_tslm3,
              test = test_ts)
```

```{r}
accuracy(fc_tslm3, test_ts) 
```

## V: Model selection

### 1: Residuals analysis

```{r}
summary(md_tslm3)$coefficients %>% tail(1)
```
```{r}
anova(md_tslm3)
```

```{r}
final_md <- tslm(UK_ts ~ season + trend + wday + month + lag365, 
                 data = UKdaily)
```

```{r}
checkresiduals(final_md)
```

### 2: Finalizing the forecast

```{r}
UK_fc_df <- data.frame(date = seq.Date(from = max(UKdaily$TIMESTAMP) + days(1), 
                                       by = "day", 
                                       length.out = h))

UK_fc_df$wday <- factor(lubridate::wday(UK_fc_df$date, label = TRUE), ordered = FALSE)

UK_fc_df$month <- factor(month(UK_fc_df$date, label = TRUE), ordered = FALSE)

UK_fc_df$lag365 <- tail(UKdaily$ND, h)

UKgrid_fc <- forecast(final_md, h = h, newdata = UK_fc_df)

```

```{r}
plot_forecast(UKgrid_fc,
              title = "The UK National Demand for Electricity Forecast",
              Ytitle = "MW",
              Xtitle = "Year")
```

