---
title: "MODELOS PREDICTIVOS (SECCION 11)"
output: html_document
date: "2024-04-29"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# CHAPTER 11: Forecasting with ARIMA Models

## I: The stationary process

### 1

```{r}
set.seed(12345)

stationary_ts <- arima.sim(model = list(order = c(1,0,0),
                                        ar = 0.5 ),
                           n = 500)
```

```{r}
library(TSstudio)

ts_plot(stationary_ts, 
        title = "Stationary Time Series",
        Ytitle = "Value",
        Xtitle = "Index")
```

### 2

```{r}
set.seed(12345)

non_stationary_ts <- arima.sim(model = list(order = c(1,1,0),ar = 0.3 ),n = 500)

ts_plot(non_stationary_ts, 
        title = "Non-Stationary Time Series",
        Ytitle = "Value",
        Xtitle = "Index")
```


### 3

```{r}
data(AirPassengers)

ts_plot(AirPassengers,
        title = "Monthly Airline Passenger Numbers 1949-1960",
        Ytitle = "Thousands of Passengers",
        Xtitle = "Year")
```

### 4: Transforming a non-stationary series into a stationary series


##### a) : Differencing time series

```{r}
ts_plot(diff(AirPassengers, lag = 1),
        title = "AirPassengers Series - First Differencing",
        Xtitle = "Year",
        Ytitle = "Differencing of Thousands of Passengers")
```

#### b)

```{r}
ts_plot(diff(diff(AirPassengers, lag = 1), 12),
        title = "AirPassengers Series - First and Seasonal Differencing",
        Xtitle = "Year",
        Ytitle = "Differencing of Thousands of Passengers")
```

#### c) : Log transformation

```{r}
ts_plot(diff(log(AirPassengers), lag = 1),
        title = "AirPassengers Series - First Differencing with Log Transformation",
        Xtitle = "Year",
        Ytitle = "Differencing/Log of Thousands of Passengers")
```

### 5: The random walk process (1)

```{r}
set.seed(12345)

library(plotly)

p1 <- plot_ly()
p2 <- plot_ly()

for(i in 1:20){
  rm <- NULL
  rw <- arima.sim(model = list(order = c(0, 1, 0)), n = 500)
  p1 <- p1 %>% add_lines(x = time(rw), y = as.numeric(rw))
  p2 <- p2 %>% add_lines(x = time(diff(rw)), y = as.numeric(diff(rw)))
}
```
```{r}
p1 %>% layout(title = "Simulate Random Walk",
              yaxis = list(title = "Value"),
              xaxis = list(title = "Index")) %>%
  hide_legend()
```
### 6: The random walk process (2)

```{r}
p2 %>% layout(title = "Simulate Random Walk with First-Order Differencing",
              yaxis = list(title = "Value"),
              xaxis = list(title = "Index")) %>%
  hide_legend()
```

## II: The AR process

### 1

```{r}
set.seed(12345)

ar2 <- arima.sim(model = list(order = c(2,0,0),
                              ar = c(0.9, -0.3)),
                 n = 500) 
```

```{r}
ts_plot(ar2, 
        title = "Simulate AR(2) Series",
        Ytitle = "Value",
        Xtitle = "Index")
```

```{r}
md_ar <- ar(ar2)
md_ar
```
### 2: Identifying the AR process and its characteristics

```{r}
par(mfrow=c(1,2))
acf(ar2)
pacf(ar2)
```


## III: The moving average process

### 1

```{r}
set.seed(12345)

ma2 <- arima.sim(model = list(order = c(0, 0, 2),
                              ma = c(0.5, -0.3)),
                 n = 500) 
```

```{r}
ts_plot(ma2, 
        title = "Simulate MA(2) Series",
        Ytitle = "Value",
        Xtitle = "Index")
```

```{r}
md_ma <- arima(ma2, order = c(0,0,2), method = "ML")
md_ma
```

### 2: Idenfitying the MA process and its characteristics

```{r}
par(mfrow=c(1,2))
acf(ma2)
pacf(ma2)
```

## IV: The ARMA model

### 1

```{r}
set.seed(12345)

arma <- arima.sim(model = list(order(1,0,2), ar = c(0.7), ma = c(0.5,-0.3)), n = 500)
```

```{r}
ts_plot(arma, 
        title = "Simulate ARMA(1,2) Series",
        Ytitle = "Value",
        Xtitle = "Index")
```

```{r}
arma_md <- arima(arma, order = c(1,0,2))
arma_md
```


### 2: Identifying an ARMA process

```{r}
par(mfrow=c(1,2))
acf(arma)
pacf(arma)
```


### 3: Manual tuning of the ARMA model

```{r}
arima(arma, order = c(1, 0, 1), include.mean = FALSE)
```

```{r}
arima(arma, order = c(2, 0, 1), include.mean = FALSE)
```

```{r}
arima(arma, order = c(1, 0, 2), include.mean = FALSE)
```

```{r}
arima(arma, order = c(2, 0, 2), include.mean = FALSE)
```

```{r}
library(forecast)
best_arma <- arima(arma, order = c(1, 0, 2), include.mean = FALSE)

checkresiduals(best_arma)
```

## V: Forecasting AR, MA, and ARMA models

```{r}
ar_fc <- forecast(md_ar, h = 100)

plot_forecast(ar_fc, 
              title = "Forecast AR(2) Model",
              Ytitle = "Value",
              Xtitle = "Year")
```

## VI: The ARIMA model

### 1: Identifying the model degree of differencing

```{r}
data(Coffee_Prices)

robusta_price <- window(Coffee_Prices[,1], start = c(2000, 1))
```

```{r}
ts_plot(robusta_price,
        title = "The Robusta Coffee Monthly Prices",
        Ytitle = "Price in USD",
        Xtitle = "Year") 
```

```{r}
acf(robusta_price)

```

```{r}
robusta_price_d1 <- diff(robusta_price)
```

```{r}
par(mfrow=c(1,2))
acf(robusta_price_d1)
pacf(robusta_price_d1)
```

```{r}
robusta_md <- arima(robusta_price, order = c(1, 1, 0))
```

```{r}
summary(robusta_md)
```

```{r}
checkresiduals(robusta_md)
```

## VII: The Seasonal ARIMA model (SARIMA)

### 1: Forecasting US monthly natural gas consumption with the SARIMA model - a case study

```{r}
data(USgas)

ts_plot(USgas,
        title = "US Monthly Natural Gas consumption", 
        Ytitle = "Billion Cubic Feet",
        Xtitle = "Year")
```

#### 1

```{r}
USgas_split <- ts_split(USgas, sample.out = 12)

train <- USgas_split$train
test <- USgas_split$test
```

```{r}
par(mfrow=c(1,2))
acf(train, lag.max = 60)
pacf(train, lag.max = 60)
```

```{r}
USgas_d12 <- diff(train, 12)

ts_plot(USgas_d12,
        title = "US Monthly Natural Gas consumption - First Seasonal Difference", 
        Ytitle = "Billion Cubic Feet (First Difference)",
        Xtitle = "Year")
```

#### 2

```{r}
USgas_d12_1 <- diff(diff(USgas_d12, 1))

ts_plot(USgas_d12_1,
        title = "US Monthly Natural Gas consumption - First Seasonal and Non-Seasonal Differencing", 
        Ytitle = "Billion Cubic Feet (Difference)",
        Xtitle = "Year")
```


```{r}
par(mfrow=c(1,2))
acf(USgas_d12_1, lag.max = 60)
pacf(USgas_d12_1, lag.max = 60)

```

#### 3

```{r}
p <- q <- P <- Q <- 0:2
```

```{r}
arima_grid <- expand.grid(p,q,P,Q)
names(arima_grid) <- c("p", "q", "P", "Q")
arima_grid$d <- 1
arima_grid$D <- 1
```

```{r}
arima_grid$k <- rowSums(arima_grid)
```

```{r}
library(dplyr)

arima_grid <- arima_grid %>% filter(k <= 7)
```

```{r}
# arima_search <- lapply(1:nrow(arima_grid), function(i){
#   md <- NULL
#   md <- arima(train, order = c(arima_grid$p[i], 1, arima_grid$q[i]), seasonal = list(order = c(arima_grid$P[i], 1, arima_grid$Q[i])))
#   
#   results <- data.frame(p = arima_grid$p[i], d = 1, q = arima_grid$q[i],
#                         P = arima_grid$P[i], D = 1, Q = arima_grid$Q[i],
#                         AIC = md$aic)
# }) %>% bind_rows() %>% arrange(AIC)

```

```{r}
#head(arima_search)
```

#### 4

```{r}
USgas_best_md <- arima(train, order = c(1,1,1), seasonal = list(order = c(2,1,1)))
USgas_best_md
```

```{r}
USgas_test_fc <- forecast(USgas_best_md, h = 12)
```

```{r}
accuracy(USgas_test_fc, test)
```

```{r}
test_forecast(USgas, 
              forecast.obj = USgas_test_fc,
              test = test)
```

#### 5

```{r}
final_md <- arima(USgas, order = c(1,1,1), seasonal = list(order = c(2,1,1)))
checkresiduals(final_md)
```

```{r}
USgas_fc <- forecast(final_md, h = 12)
plot_forecast(USgas_fc,
              title = "US Natural Gas Consumption - Forecast",
              Ytitle = "Billion Cubic Feet",
              Xtitle = "Year")
```


## VIII: The auto.arima function

### 1

```{r}
USgas_auto_md1 <- auto.arima(train)
USgas_auto_md1
```

```{r}
USgas_auto_md2 <- auto.arima(train, 
                             max.order = 5,
                             D = 1,
                             d = 1,
                             ic = "aic",
                             stepwise = FALSE, 
                             approximation = FALSE)
USgas_auto_md2
```

### 2: Linear regression with ARIMA errors & Violation of white noise assumption

```{r}
df <- ts_to_prophet(AirPassengers) %>% setNames(c("date", "y"))

df$lag12 <- dplyr::lag(df$y, n = 12)

library(lubridate)

df$month <- factor(month(df$date, label = TRUE), ordered = FALSE)

df$trend <- 1:nrow(df)
```

```{r}
par <- ts_split(ts.obj = AirPassengers, sample.out = 12)

train <- par$train

test <- par$test
```

```{r}
train_df <- df[1:(nrow(df) - 12), ]
test_df <- df[(nrow(df) - 12 + 1):nrow(df), ]
```

```{r}
md1 <- tslm(train ~ season + trend + lag12, data = train_df)
```

```{r}
checkresiduals(md1)
```

### 3: Modeling the residuals with the ARIMA model

```{r}
md2 <- auto.arima(train, 
                  xreg = cbind(model.matrix(~ month,train_df)[,-1], 
                               train_df$trend, 
                               train_df$lag12), 
                  seasonal = TRUE, 
                  stepwise = FALSE, 
                  approximation = FALSE)
```

```{r}
summary(md2)
```

```{r}
checkresiduals(md2)
```

```{r}
fc1 <- forecast(md1, newdata = test_df)

fc2 <- forecast(md2, xreg = cbind(model.matrix(~ month,test_df)[,-1], 
                                  test_df$trend, 
                                  test_df$lag12))

accuracy(fc1, test)
accuracy(fc2, test)
```

